---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# tidystats

<!-- badges: start -->
```{r}
#| echo: false
#| results: hide
#| message: false
library(badger)
```

```{r}
#| echo: false
#| results: asis
cat(
	badge_cran_release("tidystats", "green"),
	badge_cran_download("tidystats", "grand-total", "green"),
	badge_cran_download("tidystats", "last-month", "green"),
	badge_cran_download("tidystats", "last-week", "green"),
  badge_doi("10.5281/zenodo.4041859", "blue")
)
```
<!-- badges: end -->

**Author:** [Willem Sleegers](https://www.willemsleegers.com/)
**License:** [MIT](https://opensource.org/licenses/MIT)

`tidystats` is an R package aimed at sharing the output of statistical models. `tidystats` extracts the output (e.g., *t*-values, *p*-values) of statistical models (e.g., *t*-tests, regression models) and combines them into a structured file. This file can then be used to report the statistics in a manuscript or shared with others so that they can extract the statistics (e.g., for meta-analyses).

Please see below for instructions on how to install and use this package. **Do note that the package is currently in development. This means the package may contain bugs and is subject to significant changes.** If you find any bugs or if you have any feedback, please let me know by creating an issue here on Github (it's really easy to do!).

## Installation

`tidystats` can be installed from CRAN and the latest version can be installed from Github using [devtools](https://github.com/r-lib/devtools). 

```{r}
#| eval: false
devtools::install_github("willemsleegers/tidystats")
```

## Setup

Load the package and start by creating an empty list to store the results of statistical models. You can name the list whatever you want (in the example below I create an empty list called `results`).

```{r}
#| warning: false
#| message: false
library(tidystats)

results <- list()
```

## Usage

The main function is `add_stats()`. The function has 2 necessary arguments:

- `results`: The list you want to add the statistical output to.
- `output`: The output of a statistical test you want to add to the list (e.g., the output of `t.test()` or `lm()`)

Optionally you can also specify an `identifier`, the `type` of analysis, whether the analysis was `preregistered`, and/or additional `notes`.  

The `identifier` is used to identify the model (e.g., 'weight_height_correlation'). If you do not provide one, one is automatically created for you.

The `type` argument specifies the type of analysis as primary, secondary, or exploratory.

The `preregistered` argument is used to indicate whether the analysis was preregistered or not.

Finally the `notes` argument is used to add additional information which you may find fruitful.

## Supported statistical functions

**Package:** stats

- `t.test()`
- `cor.test()`
- `chisq.test()`
- `wilcox.test()`
- `fisher.test()`
- `oneway.test()`
- `aov()`
- `lm()`
- `anova()`

**Package:** lme4

- `lmer()`

**Package:** lmerTest

- `lmer()`

**Package:** BayesFactor

- `generalTestBF()`
- `lmBF()`
- `regressionBF()`
- `ttestBF()`
- `anovaBF()`
- `correlationBF()`
- `contingencyTableBF()`
- `proportionBF()`
- `meta.ttestBF()`

**Package:** tidystats

- `describe_data()`
- `count_data()`

## Example

```{r}
#| echo: false
#| message: false
library(dplyr)
```

In the following example we perform several tests, add them to a list, and save the list to a file.

```{r}
#| eval: false
# Conduct three different analyses
# t-test:
sleep_test <- t.test(extra ~ group, data = sleep, paired = TRUE)

# lm:
ctl <- c(4.17,5.58,5.18,6.11,4.50,4.61,5.17,4.53,5.33,5.14)
trt <- c(4.81,4.17,4.41,3.59,5.87,3.83,6.03,4.89,4.32,4.69)
group <- gl(2, 10, 20, labels = c("Ctl","Trt"))
weight <- c(ctl, trt)
lm_D9 <- lm(weight ~ group)

# ANOVA:
npk_aov <- aov(yield ~ block + N*P*K, npk)

# Create an empty list
results <- list()

# Add the analyses to the empty list
results <- results %>%
  add_stats(sleep_test, type = "primary") %>%
  add_stats(lm_D9, preregistered = FALSE) %>%
  add_stats(npk_aov, notes = "An ANOVA example")

# Save the results to a file
write_stats(results, "results.json")
```

This results in a .json file that contains all the statistics from the three models. If you want to see what this file looks like, you can inspect it [here](https://github.com/WillemSleegers/tidystats/blob/master/inst/results.json).

## Reporting statistics

If you want to report the statistics in a manuscript, you can do so with a Word add-in that is currently in development [here](https://github.com/WillemSleegers/tidystats-Word-add-in).

## More resources

Do you have questions or comments about `tidystats`? Create an issue here on Github or contact me via [Twitter](https://twitter.com/willemsleegers).