
<!-- README.md is generated from README.Rmd. Please edit that file -->

# tidystats

<!-- badges: start -->
```{r}
#| echo: false
#| results: hide
#| message: false
library(badger)
```

```{r}
#| echo: false
#| results: asis
cat(
  badge_cran_release("tidystats", "green"),
  badge_cran_download("tidystats", "grand-total", "green"),
  badge_cran_download("tidystats", "last-month", "green"),
  badge_cran_download("tidystats", "last-week", "green"),
  badge_doi("10.5281/zenodo.4041859", "blue")
)
```
<!-- badges: end -->

**Author:** [Willem Sleegers](https://www.willemsleegers.com/)
**License:** [MIT](https://opensource.org/licenses/MIT)

tidystats is an R package for sharing and reporting statistics. tidystats 
extracts statistics from the output of statistical functions 
(e.g., `t.test()`, `lm()`) and stores them in a structured format. The resulting
file can be shared with others and used in popular text editors to reproducibly 
report the statistics.

Please see below for instructions on how to install and use this package. 

**Do note that the package is currently in development. This means the package 
may contain bugs and is subject to significant changes.** If you find any bugs 
or if you have any feedback, please let me know by creating an issue here on 
Github (it's really easy to do!).

## Installation

tidystats can be installed from CRAN.

```{r}
#| eval: false
install.packages("tidystats")
```

You can also install the development version from GitHub using the 
[`remotes`](https://github.com/r-lib/remotes) package.

```{r}
#| eval: false
remotes::install_github("willemsleegers/tidystats")
```

## Usage

The main function is `add_stats()`. The function has 2 necessary arguments:

- `list`: The list you want to add the statistical output to. 
- `output`: The output of a statistics function you want to add to the list 
  (e.g., the output of `t.test()` or `lm()`)

Optionally, you can also specify meta-information: 

- `identifier`: A unique identifier to identify the model 
  (e.g., 'weight_height_correlation'). If you do not provide one, one is 
  automatically created for you.
- `type`: A type that specifies the type of analysis as primary, secondary, or 
  exploratory.
- `preregistered`: Whether the analysis was preregistered or not.
- `notes`: Additional information you think is useful to record.

Once you have added all statistics to a list, you can write the contents to a 
file using the `write_stats()` function.

## Example

The following example shows how to combine and save the statistics from three
different statistical tests.

```{r}
#| eval: false
# Conduct a t-test, regression, and an ANOVA
sleep_test <- t.test(extra ~ group, data = sleep, paired = TRUE)

ctl <- c(4.17, 5.58, 5.18, 6.11, 4.50, 4.61, 5.17, 4.53, 5.33, 5.14)
trt <- c(4.81, 4.17, 4.41, 3.59, 5.87, 3.83, 6.03, 4.89, 4.32, 4.69)
group <- gl(2, 10, 20, labels = c("Ctl", "Trt"))
weight <- c(ctl, trt)
lm_D9 <- lm(weight ~ group)

npk_aov <- aov(yield ~ block + N * P * K, npk)

# Create an empty list to add the statistics to
statistics <- list()

# Add the statistics, as well as some meta-information
statistics <- statistics |>
  add_stats(sleep_test, type = "primary") |>
  add_stats(lm_D9, preregistered = FALSE) |>
  add_stats(npk_aov, notes = "An ANOVA example")

# Save the statistics to a file
write_stats(statistics, "statistics.json")
```

The result is a .json file that contains all the statistics from the three 
statistical tests. If you want to see what this file looks like, you can inspect
it [here](https://github.com/WillemSleegers/tidystats/blob/master/tests/data/main.json).

## Supported statistical functions

tidystats supports functions from the following packages:

- `stats`
- `lme4`
- `lmerTest`
- `BayesFactor`
- `tidystats`

In some cases you need to provide extra information to the `add_stats()` 
function in order for tidystats to correctly extract the statistics. You can 
see a list of functions that require the `class` argument in the documentation
of the `add_stats()` function (`?add_stats`).

If you want to use tidystats on an unsupported function, there are two things
you can do:

1. Request support for the new function by creating an Issue. For more 
   information on what it means to add support for a function, see the 
   [tidystats development](https://willemsleegers.github.io/tidystats/articles/tidystats-development.html) 
   vignette.
2. Manually extract the statistics and add them via `add_stats()` using the 
   `custom_stats()` function. See the vignette on adding 
   [custom statistics](https://willemsleegers.github.io/tidystats/articles/custom-statistics.html)
   for more information.

## Reporting statistics

The file created with the `write_stats()` function can be used in several text
editor add-ins to reproducibly report the statistics. For more information on 
these add-ins, please see the [tidystats website](https://www.tidystats.io) or 
their GitHub pages:

- [Microsoft Word add-in](https://github.com/WillemSleegers/tidystats-Word-add-in)
- [Google Docs add-in](https://github.com/WillemSleegers/tidystats-Google-Docs-add-in)

## More information

See the [tidystats website](https://www.tidystats.io) for more information, 
such as a FAQ, tips and tricks, as well as how to receive (and give) support. 

If you have any questions or comments, feel free to create an 
[issue](https://github.com/WillemSleegers/tidystats/issues) here on GitHub or 
see the website for ways to contact me.