---
title: "Introduction to tidystats"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction-to-tidystats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Why use tidystats? 

The tidystats package is designed to address two problems common in scientific research: incomplete and incorrect statistics reporting. The problem of incomplete statistics reporting likely stems from a fundamental trade-off between wanting to be comprehensive on the one hand, and providing a clear narrative on the other. The problem of incorrect statistics reporting is likely caused by manually copy-pasting statistical output from the output window into a text editor. tidystats addressed these two problems by enabling researchers to combine their statistical analyses into a single file, which can subsequently be used for statistical reporting using a Microsoft Word add-in. 

## How to use tidystats?

```{r setup}
library(tidystats)
```
