---
title: "Read and use a tidystats file"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Read and use a tidystats file}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Besides using `tidystats` in combination with a text editor add-in to report statistics, you can also use `tidystats` to convert a list of statistics into a data frame. This enables researchers to then easily extract specific statistics to perform additional analyses with (e.g., meta-analyses). Below is an example of how to convert a list of statistics to a data frame and several simple operations.

```{r example1_eval, eval = FALSE, message = FALSE}
library(tidystats)
library(dplyr)

statistics <- read_stats("statistics.json") |>
  tidy_stats_to_data_frame()

p_values <- filter(statistics, statistic == "p")
```

With the current example, this results in the following data frame:

```{r example1_no_eval, message = FALSE, echo = FALSE}
library(tidystats)
library(dplyr)
library(knitr)

statistics <- read_stats("statistics.json") |>
  tidy_stats_to_data_frame()

p_values <- filter(statistics, statistic == "p")

options(knitr.kable.NA = '')

p_values |>
  kable(format = "markdown")
```

Alternatively, you can select all the significant p-values:

```{r example2_no_eval, eval = FALSE}
sig_p_values <- filter(statistics, statistic == "p" & value < .05)
```

```{r example2_eval, echo = FALSE}
sig_p_values <- filter(statistics, statistic == "p" & value < .05)

sig_p_values |>
  kable(format = "markdown")
```

This could be useful if you want to conduct a [p-curve analysis](https://journals.sagepub.com/doi/abs/10.1177/1745691614553988). Although do note that you should not blindly select all *p*-values. You should select only the *p*-values that are relevant to a particular hypothesis. If researchers provide the correct meta-information for each test (e.g., by indicating whether it is a primary analysis), this could help meta-researchers make correct decisions about which statistics to include in their analyses.

In short, by reading a `tidystats`-produced file of statistics, you can convert the statistics to a data frame using the `tidy_stats_to_data_frame` function and apply common data transformation functions to extract specific statistics.
